\chapter{Future Work}
The scope of this project is rather broad and the size of the project is quite large given the limited timeframe. For this and other reasons mentioned throughout the report the state of the system is far from being perfect, yet it is a functional prototype of the envisioned Equine Health Monitor. 

We would like to use this chapter to discuss solutions to problems in the system design and implementation that we discovered, and propose a number of improvements that can be made to the system to increase its usability and extend the possible applications.


\section{Solutions for design problems}
During the testing phase of the wireless communication module of our system we realized, that the chosen solution is not able to deliver the bandwidth that we require to transfer audio recording from the monitoring devices to the base station in a reasonable amount of time. 

The problem is that the ZigBee protocol is designed for low bandwidth requirements, for even it is a low power protocol, the energy cost per transmitted byte is a lot higher than for other wireless solutions, such as 802.11 or Bluetooth.

Because we do not require a constant connection, but are able transmit data periodically in a burst transmit style, it is feasible to use an alternative solution and run it at a low duty cycle. 

To be more specific in terms of our design, there are pin compatible replacements for XBee devices that implement the 802.11 standard and the manufacturer of the XBee devices is working on a 802.11 based XBee version as well (currently available as a development kit, not retail)

Replacing the XBees with alternative devices doesn’t require modifications to the system design, and because of the Object Oriented design of the software system and well defined interfaces the changes that are required in the software stack are limited to a clearly defined scope.

An audio compression scheme is then recommended if the system is required to perform audio recordings often and/or send this data over the wireless connection. Such a scheme was not implemented in the project mainly the short amount of time available for development and that a deeper knowledge of the signals is required but could be implemented in an improved version of the system.

\section{Possible improvements}
The device is designed as an equine health monitor, however the need for a health monitor is not only limited to horses. As there are almost no horse specific parts used in the system, the device could be adapted for the use on human subjects or agents of other mammal species. The only adaptation that has to be made is the usage of a target specific heart rate monitor. 

The scope of the project is rather large. It is possible improve some of the features to evolve it into a more specialized system. One possible solution could be a tracking system using GPS and accelerometer data. These data can also be used for health-related applications such as tracking the movement of the animals and detecting extraordinary situations. 

Another field of application could be professional horse training, to monitor the heart rate in relation to the covered distance over an average of the velocity which allows to analyze the improvements of a horse’s stamina over time.

As the system works as a health profiler, it is possible extend it by adding signal analysis features to perform autonomous diagnosis or issue warnings when there is a problem with the vital signs. Further research has to be undertaken to determine relations of parameters and indicators for certain diseases.

In the current implementation, the gut sound monitor works periodically. The problem with this approach is introduced noise when the horse is moving. One way to improve the behavior is to disabling the audio recording conditionally, if the horse is moving and delay it until the horse stays at a stable position for a period of time. The (already available) accelerometer and GPS data can be used to implement this feature.

In the current state of the system the webinterface is not very user friendly when it comes to analyzing large amounts of collected data. This could be improved by using advances web interface technologies like Ajax to build dynamic version of the website that behaves more like a desktop application than a website.

As accessibility is an important goal of the project it would be a useful addition to extend the website so that a handheld optimized version of the website is delivered when the user accesses the base station with a smartphone.

The limitations of the used XBee devices was discussed in section \ref{sec:wireless_testing}. To be able to transmit audio recordings the ZigBee network should be replaced with a faster wireless connection. Pin compatible XBee replacement ICs based on the 802.11 standard exist and require only a modification on the software part of the system.

Even though the system was designed with energy efficiency in mind from the very beginning, there are subsystems that require further optimization for low energy consumption. The XBee devices are not sent to sleep in the current system, because the delay before they manage to reestablish a connection is too long.
